---
title: "NNIP Data Presentation"
author: "Cristina Martinez, MPH - CI:Now"
date: "July 28, 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Downloading Census Data using Tidycensus

You will need **Census API Key** to use tidycensus
You can get one here: <https://api.census.gov/data/key_signup.html>


Resources used to put this demonstration together:

Analyzing US Census Data by Kyle Walker (Book):
<https://walker-data.com/census-r/index.html>

Basic usage of Tidycensus:
<https://walker-data.com/tidycensus/articles/basic-usage.html>

Working with Census microdata:
<https://walker-data.com/tidycensus/articles/pums-data.html>

```{r}
#install.packages("tidycensus") ## To install package. Only has to be done once.

#Load Tidycensus library
library(tidycensus)
library(tidyverse) #another useful package for manipulating and reshaping data

## INSTALL KEY AND STORE IN R ENVIRONMENT
census_api_key("YOUR KEY GOES HERE", install = TRUE)
```

# Tidycensus has the following core functions
```{r}
get_decennial() # US Decennial Census API for 2000, 2010, 2020

get_acs()       # 1 (2005) & 5 (2005-2009) Year ACS

get_estimates() # Population Estimates by state, county, and MSA

get_pums()      # ACS Public Use Microdata Samples

get_flows()     # ACS Migration Flows
```

We'll focus on get_acs() and get_pums() for this demonstration.

# Pulling ACS data
```{r}
# SEARCH FOR VARIABLES

# detailed tables
load_variables(2020, "acs5", cache = TRUE) # can also call "acs1"

#Save to an object
var20 <- load_variables(2020, "acs5", cache = TRUE) # can also call "acs1"
View(var20)

# subject tables
load_variables(2020, "acs5/subject", cache = TRUE)

# profile tables
load_variables(2020, "acs5/profile", cache = TRUE)
```

# Pull Median Household Income
```{r}
tx <- get_acs(geography = "county", # can also call "us", "region", "state", "tract", "block group" among many other geos
              variables = "B19013_001", 
              state = "TX", 
              year = 2020,
              survey = "acs5")

tx
```

# Pull Median Household Income for all households and Hispanic/Latino households
```{r}
tx <- get_acs(geography = "county", 
              variables = c("B19013_001", "B19013I_001"), #pull multiple variables
              output = "wide", # can change the output of the data so it's wide
              state = "TX", 
              year = 2019, # change year for different survey years
              survey = "acs1") #change survey for acs1 or acs5

tx
```

# You can also pull an entire table by using "table = " instead of "variable = "
```{r}
tx <- get_acs(geography = "county", 
              table = "B01001", # Population by age and sex. Can only pull one table at a time
              output = "wide", 
              state = "TX", 
              year = 2020,
              survey = "acs5") 

tx
```

# Easily calculate new estimates and margins of error
```{r}
tx <- get_acs(geography = "tract",
              variables = c(totalpop = "B01001_001", malepop = "B01001_002", femalepop = "B01001_026"),
              output = "wide",
              county = "Bexar",
              state = "TX") #defaults to 2020 ACS 5 year

tx

tx <- tx %>%
  mutate(
    percent.male = malepopE/totalpopE * 100, 
    percent.male.moe = moe_prop(malepopE, totalpopE, malepopM, totalpopM) * 100, # Calculates the moe for the new proportion
    percent.female = femalepopE/totalpopE * 100, 
    percent.female.moe = moe_prop(femalepopE, totalpopE, femalepopM, totalpopM) * 100
  ) %>%
  select(NAME, percent.male:percent.female.moe) %>%
  arrange(NAME) #Arranges data frame in ascending order based on the NAME column

tx
```

# Pulling data for multiple years
```{r}
years <- 2017:2019
names(years) <- years

pop.by.age <- map_dfr(years, ~{
  get_acs(
    geography = "county",
    variables = c("B01001_001", "B01001_002", "B01001_026"),
    state = "TX",
    output = "wide",
    survey = "acs1",
    year = .x
  )
}, .id = "year")

pop.by.age
```

# Mapping using Tidycensus
```{r}
options(tigris_use_cache = TRUE) # used to cache the downloaded geographic data on the user’s computer

dc_income <- get_acs(
  geography = "tract", 
  variables = "B19013_001", #median household income
  state = "DC", 
  year = 2020,
  geometry = TRUE # use this argument when you want to download corresponding shape files
)

dc_income

plot(dc_income["estimate"])

# See Chapter 6 of Analyzing US Census Data for more detailed examples
```

# Working with microdata (PUMS)
```{r}
#SEARCH FOR VARIABLES
pums_vars_19 <- pums_variables %>% 
  filter(year == 2019, survey == "acs1")

View(pums_vars_19)
```

# Pull microdata
```{r}
hi_pums <- get_pums(
  variables = c("PUMA", "SEX", "AGEP", "SCHL"), # Pulling sex, age, and educational attainment
  state = "HI",
  survey = "acs5",
  year = 2020,
  recode = TRUE, # This argument will give you the data labels
  #rep_weights = "person" # Can also get "household", or "both". Downloading replicated weights takes about 5-10 minutes.
  )

hi_pums
#Automatically returns unique identifiers, individual and household weights, and state
```

# Estimated population
```{r}
sum(hi_pums$PWGTP) # Individuals
sum(hi_pums$WGTP)  # Households
```

# Estimated population by PUMA and SEX
```{r}
hi_pums %>% 
  count(PUMA, SEX_label, wt = PWGTP)

# See Chapters 9 and 10 of Analyzing US Census Data for more information on analyzing PUMS data
```
